---
title: "LM results"
author: "Robyn Sumner"
date: "2024-01-13"
output: html_document
---

# This code was used to explore the Manchester and Liverpool dataset and created the grpahs used in the results section 4.2

```{r}
library(tidyverse)
library(cli)
library(tidyr)
library(tidyverse)
library(cli)
library(tidyr)
library(ggplot2)
library(GGally)
library(corrgram)
library(dplyr)

all = read.csv("C:/Users/robyn/OneDrive/Documents/1 - Uni of Bristol/Y3 Geography/Dissertation/Data/Data/Tidy Data - All.csv")

emissions = all |> 
  pivot_longer(
    cols = !(site:long),
    names_to = c("date", ".value"), 
    names_sep = "_",
    values_drop_na = TRUE
  )

emissions$season =  ifelse(
  (emissions$date) %in% c("Mar.22", "April.22", "May.22"), "spring",
  ifelse(
    (emissions$date) %in% c("June.22", "July.22", "Aug.22"), "summer",
    ifelse(
      (emissions$date) %in% c("Sept.22", "Oct.22", "Nov.21"), "autumn",
      "winter"
    )
  )
)

summer_data <- emissions %>%
  filter(season == "summer")

spring_data <- emissions %>%
  filter(season == "spring")

autumn_data <- emissions %>%
  filter(season == "autumn")

winter_data <- emissions %>%
  filter(season == "winter")

emissions$Diffusive_CH4_Flux_Mean = emissions$methane.nmolm.s * 86400 * 1/1000000
```

# Correlations graph

```{r}
emissions_renamed = emissions

emissions_renamed$Methane = emissions_renamed$methane.nmolm.s
emissions_renamed$DO = emissions_renamed$DO.mg.L
emissions_renamed$ORP = emissions_renamed$ORP.mV
emissions_renamed$fDOM = emissions_renamed$fDOM.ppb
emissions_renamed$EC = emissions_renamed$CON.µS.cm
emissions_renamed$Temperature = emissions_renamed$Temp.C

jpeg(filename =  "Graph 6a.jpg", units = "cm", res = 1000, height = 20, width = 20)


ggcorr(emissions_renamed[,c(19,20,21,22,23,18)], 
       method = c("pairwise.complete.obs", "pearson"),
       nbreaks = 10,
       name = "Pearson Correlation Coefficient (r)",
       legend.size = 11,
       label = TRUE,
       label_size = 4, 
       label_color = "black",
       label_round = 2,
       layout.exp =2
#       high = "#cc3311",
#       low = "#0077bb"
        ) 

dev.off()

```

# The relationship between methane flux and the predictors

```{r}
# Selecting only the variables you want to plot
selected_vars <- c("DO.mg.L", "ORP.mV", "fDOM.ppb", "CON.µS.cm", "Temp.C")

# Subsetting the data frame to include only the selected variables
selected_data <- emissions[, c("methane.nmolm.s", selected_vars)]

# Reshaping the data into long format
selected_data_long <- tidyr::pivot_longer(selected_data, cols = -methane.nmolm.s, names_to = "variable", values_to = "value")

# Create a function to rename the variables
rename_variable <- function(variable) {
  if(variable == "DO.mg.L") {
    return("Dissolved Oxygen log(mg/L)")
  } else if(variable == "ORP.mV") {
    return("ORP log(mV)")
  } else if(variable == "fDOM.ppb") {
    return("fDOM log(ppb)")
  } else if(variable == "CON.µS.cm") {
    return("Electrical Conductivity log(µS/cm)")
  } else if(variable == "Temp.C") {
    return("Temperature log(°C)")
  } else {
    return(variable)
  }
}

# Apply the function to rename the variables
selected_data_long$variable <- sapply(selected_data_long$variable, rename_variable)

jpeg(filename =  "Graph 7.jpg", units = "cm", res = 1000, height = 20, width = 20)

# Plotting
ggplot(selected_data_long, aes(x = value, y = methane.nmolm.s)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +  
  geom_smooth(method = "lm", se = FALSE, show.legend = TRUE) +
  facet_wrap(~ variable, scales = 'free_x') +
  ylab(expression(Methane ~ log(nmol ~ m^2 ~ s^1))) +  
  theme(axis.title.x = element_blank())

dev.off()
```

# Box and whisker plot of the dates/seasons against the methane flux

```{r}
# Load required libraries
library(ggplot2)
library(patchwork)

# Create a custom order for the dates
custom_order <- c("Dec.21", "Jan.22", "Feb.22", "Mar.22", "April.22", "May.22", "June.22", "July.22", "Aug.22","Sept.22", "Oct.22","Nov.21")

# Convert the 'date' column in your data frame to a factor with the custom order
emissions$date <- factor(emissions$date, levels = custom_order)

# Plot 1: Violin plot with custom order of dates
violin <- ggplot(emissions, aes(x = date, y = Diffusive_CH4_Flux_Mean, fill = season)) +
  geom_violin(alpha = 0.6) + 
  scale_y_log10() +  
  theme_minimal() +
  scale_x_discrete(labels = c("Sept.22" = "September", 
                              "Oct.22" = "October", 
                              "Nov.21" = "November", 
                              "Dec.21" = "December", 
                              "Jan.22" = "January", 
                              "Feb.22" = "February", 
                              "Mar.22" = "March", 
                              "April.22" = "April", 
                              "May.22" = "May", 
                              "June.22" = "June", 
                              "July.22" = "July", 
                              "Aug.22" = "August")) +
  scale_fill_manual(values = c("spring" = "#EE6677", 
                                "summer" = "#CCBB44", 
                                "autumn" = "#229933", 
                                "winter" = "#4477AA"),
                    name = "Seasons") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
           text = element_text(size = 12, family = "sans")) +
  labs(x = NULL, y = expression(Methane ~ predictions ~ log(nmol ~ m^2 ~ s^1)))

# Create a custom order for the dates
custom_order2 <- c("winter", "spring", "summer", "autumn")

# Convert the 'date' column in your data frame to a factor with the custom order
emissions$season <- factor(emissions$season, levels = custom_order2)


# Plot 2: Violin plot with custom order of seasons
violin2 <- ggplot(emissions, aes(x = season, y = Diffusive_CH4_Flux_Mean, fill = season)) +
  geom_violin(alpha = 0.6) + 
  scale_y_log10() +
  theme_minimal() +
  scale_x_discrete(labels = c("spring" = "Spring", 
                              "summer" = "Summer", 
                              "autumn" = "Autumn", 
                              "winter" = "Winter")) +
  scale_fill_manual(values = c("spring" = "#EE6677", 
                               "summer" = "#CCBB44", 
                               "autumn" = "#229933", 
                               "winter" = "#4477AA"),
                    name = "Seasons") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_blank(),  # Remove y-axis text
        axis.ticks.y = element_blank(),  # Remove y-axis ticks
        text = element_text(size = 12, family = "sans")) +
  labs(x = NULL, y = NULL)  # Remove y-axis label

violina = violin + geom_boxplot(width=0.1)
violinb = violin2 + geom_boxplot(width=0.1)

# Arrange plots side by side with shared y-axis and legend
violin_combined <- violina + violinb +
  plot_layout(widths = c(3, 1), guides = 'collect')

jpeg(filename =  "Graph 8.jpg", units = "cm", res = 1000, height = 20, width = 20)

# Print the combined plot
violin_combined

dev.off()
```

# Box and whisker plot of temperature and methane flux

```{r}
# Create a custom order for the dates
custom_order2 <- c("winter", "spring", "summer", "autumn")

# Convert the 'date' column in your data frame to a factor with the custom order
emissions$season <- factor(emissions$season, levels = custom_order2)


# Plot 2: Violin plot with custom order of seasons
violin2 <- ggplot(emissions, aes(x = season, y = Temp.C, fill = season)) +
  geom_violin(alpha = 0.6) + 
  scale_y_log10() +
  theme_minimal() +
  scale_x_discrete(labels = c("spring" = "Spring", 
                              "summer" = "Summer", 
                              "autumn" = "Autumn", 
                              "winter" = "Winter")) +
  scale_fill_manual(values = c("spring" = "#EE6677", 
                               "summer" = "#CCBB44", 
                               "autumn" = "#229933", 
                               "winter" = "#4477AA"),
                    name = "Seasons") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 14),  # Adjust size here
        axis.title.x = element_blank(),  # Remove x-axis label
        text = element_text(size = 12, family = "sans")) +
  ylab("Temperature log(°C)")  # Adding y-axis label

violinb = violin2 + geom_boxplot(width=0.1)

jpeg(filename =  "Graph 10.jpg", units = "cm", res = 1000, height = 20, width = 20)

# Print the combined plot
violinb

dev.off()

mean(subset_spring$Temp.C)
```

# Statistics 

```{r}
emissions$log_methane.nmolm.s = log10(emissions$methane.nmolm.s)

emissions_less = emissions[-52, ]

linear_DO = lm(log_methane.nmolm.s ~ DO.mg.L, data = emissions)
linear_pH = lm(log_methane.nmolm.s ~ pH, data = emissions)
linear_ORP = lm(log_methane.nmolm.s ~ ORP.mV, data = emissions)
linear_fDOM = lm(log_methane.nmolm.s ~ fDOM.ppb, data = emissions)
linear_CON = lm(log_methane.nmolm.s ~ CON.µS.cm, data = emissions)
linear_TEMP = lm(log_methane.nmolm.s ~ Temp.C, data = emissions)
linear_DO_TEMP = lm(log_methane.nmolm.s ~ DO.mg.L + Temp.C, data = emissions)
linear_DO_CON = lm(log_methane.nmolm.s ~ DO.mg.L + CON.µS.cm, data = emissions)
linear_DO_fDOM = lm(log_methane.nmolm.s ~ DO.mg.L + fDOM.ppb, data = emissions)

summary(linear_DO)
# 21%
summary(linear_pH)
# 0.01, not significant
summary(linear_ORP)
# not significant
summary(linear_fDOM)
# not significant
summary(linear_CON)
# 16, significant
summary(linear_TEMP)
# 18%, significant
summary(linear_DO_TEMP)
# 25%, DO significant, Temp semi significant
summary(linear_DO_CON)
# 22% not significant
summary(linear_DO_fDOM)

# residuals vs predictions
emissions |> mutate(
  residual = residuals(linear_DO_TEMP),
  prediction = predict(linear_DO_TEMP)
  ) |>
ggplot(aes(x=prediction, y=residual)) + 
    scale_x_log10() +
  scale_y_log10() +
  geom_point() + geom_smooth(color='red', se=F)
# residuals q-q plot
emissions |> mutate(
  residuals = residuals(linear_DO_TEMP),
  ) |>
ggplot(aes(sample = residuals)) + 
  geom_qq() + geom_qq_line(color = "red")
```

# Are the seasons statistically different?

```{r}
# Can we use parametric?
subset_autumn = subset(emissions, season == 'autumn')
subset_winter = subset(emissions, season == 'winter')
subset_spring = subset(emissions, season == 'spring')
subset_summer = subset(emissions, season == 'summer')
# methane 
shapiro.test(x=subset_autumn$methane.nmolm.s)
shapiro.test(x=subset_winter$methane.nmolm.s)
shapiro.test(x=subset_spring$methane.nmolm.s)
shapiro.test(x=subset_summer$methane.nmolm.s)
# Not all parametric therefore will use non-parametric testing
library(broom)
kruskal.test(season~methane.nmolm.s, data = emissions)

```

The Kruskal-Wallis test is a non-parametric test used to determine whether there are differences in the distribution of a continuous variable among two or more groups.

With a p-value of 0.4847, and assuming a significance level of 0.05, you would fail to reject the null hypothesis. This suggests that there is not enough evidence to conclude that there are significant differences in the distribution of methane.nmolm.s among the different seasons.

In other words, the test did not find a statistically significant difference in the levels of methane emissions across the seasons.

```{r}
subset_autumn = subset(emissions, season == 'autumn')
subset_winter = subset(emissions, season == 'winter')
subset_spring = subset(emissions, season == 'spring')
subset_summer = subset(emissions, season == 'summer')

wilcox.test(subset_autumn$methane.nmolm.s, subset_winter$methane.nmolm.s)
wilcox.test(subset_autumn$methane.nmolm.s, subset_spring$methane.nmolm.s)
wilcox.test(subset_autumn$methane.nmolm.s, subset_summer$methane.nmolm.s)


wilcox.test(subset_winter$methane.nmolm.s, subset_spring$methane.nmolm.s)
wilcox.test(subset_winter$methane.nmolm.s, subset_summer$methane.nmolm.s)

wilcox.test(subset_spring$methane.nmolm.s, subset_summer$methane.nmolm.s)
```

autumn different to winter
autumn different to spring
winter different to summer
spring different to summer

# Are the tempertures of the seasons statistically different?

```{r}
# Can we use parametric?

# temperature 
shapiro.test(x=subset_autumn$Temp.C)
shapiro.test(x=subset_winter$Temp.C)
shapiro.test(x=subset_spring$Temp.C)
shapiro.test(x=subset_summer$Temp.C)
# Not all parametric therefore will use non-parametric testing
library(broom)
kruskal.test(season~Temp.C, data = emissions)

```

```{r}
subset_autumn = subset(emissions, season == 'autumn')
subset_winter = subset(emissions, season == 'winter')
subset_spring = subset(emissions, season == 'spring')
subset_summer = subset(emissions, season == 'summer')

wilcox.test(subset_autumn$Temp.C, subset_winter$Temp.C)
wilcox.test(subset_autumn$Temp.C, subset_spring$Temp.C)
wilcox.test(subset_autumn$Temp.C, subset_summer$Temp.C)


wilcox.test(subset_winter$Temp.C, subset_spring$Temp.C)
wilcox.test(subset_winter$Temp.C, subset_summer$Temp.C)

wilcox.test(subset_spring$methane.nmolm.s, subset_summer$methane.nmolm.s)
```

