---
title: "Random Forest"
author: "Robyn Sumner"
date: "2024-01-06"
output: html_document
---
```{r, warning=FALSE}
# Setting the working directory 
setwd("C:/Users/robyn/OneDrive/Documents/1 - Uni of Bristol/Y3 Geography/Dissertation/Data/Modelling")
# Loading the libraries
library(sf)
library(lubridate)
library(ggplot2)
library(corrr)
library(magrittr)
library(tidymodels) 
library(dplyr)
library(tidyr)
# Reading in the data
data = read.csv("/Users/robyn/OneDrive/Documents/1 - Uni of Bristol/Y3 Geography/Dissertation/Data/GRiMeDB.rda/data.csv")
lm_data = read.csv("/Users/robyn/OneDrive/Documents/1 - Uni of Bristol/Y3 Geography/Dissertation/Data/RiverMethaneFlux-master/RiverMethaneFlux-master/data/processed/LM - Data for Modelling.csv")
sites_df <- read.csv("/Users/robyn/OneDrive/Documents/1 - Uni of Bristol/Y3 Geography/Dissertation/Data/GRiMeDB.rda/LM_site.csv")
conc_df <- read.csv("/Users/robyn/OneDrive/Documents/1 - Uni of Bristol/Y3 Geography/Dissertation/Data/GRiMeDB.rda/LM_concentration.csv")
flux_df <- read.csv("/Users/robyn/OneDrive/Documents/1 - Uni of Bristol/Y3 Geography/Dissertation/Data/GRiMeDB.rda/LM_flux.csv")
labeller_vars <- read.csv("/Users/robyn/OneDrive/Documents/1 - Uni of Bristol/Y3 Geography/Dissertation/Data/RiverMethaneFlux-master/RiverMethaneFlux-master/data/processed/variables_names.csv") 
```

```{r}
# Take out data with NAs as they cannot be modelled
subset_data = data[-c(1,2,3,4,5,6,7,8,9,10,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,82,83,84)]

```


# Exploring the distribution of variables
# However, the data will not be transformed because random forests do not need normally distributed data

```{r, results='hide', eval=FALSE}
library(ggplot2)

ggplot(subset_data, aes(x = Diffusive_CH4_Flux_Mean)) +
  geom_histogram(fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = slop)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = popdens)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = T_OC)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = T_CACO3)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = T_CASO4)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = T_ESP)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = S_OC)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = area)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = wetland)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = N_groundwater_agri)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = N_groundwater_nat)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = N_deposition_water)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = N_retention_subgrid)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = P_aquaculture)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = P_gnpp)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = P_background)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = P_load)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = P_point)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = P_surface_runoff_agri)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = P_surface_runoff_nat)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = gw_month)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = k_month)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

ggplot(subset_data, aes(x = q_month)) +
  geom_histogram( fill = "blue", color = "black", alpha = 0.7)

```

# Exploring the correlations between variables to identify which ones to remove

```{r}
library(corrr)
library(dplyr)
# The pearson correlations for diffusive methane flux against the predictors
corr_ch4 <- subset_data %>%
  select_if(is.numeric) %>%
  correlate(method = "pearson", use = "pairwise.complete.obs") %>%
  focus(Diffusive_CH4_Flux_Mean)

# The coefficients from highest to lowest correlation
corr_ch4 %>% 
  arrange(desc(abs(Diffusive_CH4_Flux_Mean))) %>% 
  print(n = 50)
```

```{r, warning=FALSE}
# Select only the numeric columns
numeric_data <- subset_data[, sapply(subset_data, is.numeric)]
# Compute the correlation matrix using pearson
correlation_matrix <- cor(numeric_data, method = "pearson")
# Filter correlation matrix based on thresholds (e.g., 0.95 and -0.95)
filtered_correlation_matrix <- correlation_matrix * ((correlation_matrix > 0.95) | (correlation_matrix < -0.95))
```

Soil properties used also contained a value for the top and bottom soil layers, which were highly related, and so we retained only the top soil properties for modelling.
```{r}
# Select useful variables for the model, some variables were removed due to a high correlation with other ones and because of high NAs and because you cannot plot it and biome_label
variables_to_remove <- c('wetland_cover', 'biome_label','GPP_yr','gpp_month', "S_SAND", 'S_SILT', 'S_CLAY', 'S_REF_BULK_DENSITY','S_PH_H2O','S_BS','S_TEB', 'S_CACO3','S_ESP', 'S_OC', 'S_CEC_SOIL', 'S_CEC_CLAY', 'S_BULK_DENSITY', 'S_CASO4', "S_GRAVEL", 'pyearRA', "pyearRH",'N_groundwater_nat','N_load','N_point','N_aquaculture','N_gnpp','N_retention','uparea.y','runoff_yr','nee_month', 'tavg_month')

library(tidymodels) 

data_for_model <-  subset_data %>%
  select(-all_of(variables_to_remove)) %>% 
  drop_na()

```


```{r}
library(tidymodels) 

set.seed(280)

data_split <- initial_split(data_for_model, prop = 3/4)
train_data <- training(data_split)
test_data  <- testing(data_split)

data_recipe = recipe(Diffusive_CH4_Flux_Mean ~., data = train_data)
```

```{r}
tune_spec <- rand_forest(
  mtry = tune(),
  trees = tune(),
  min_n = tune()) %>%
  set_mode("regression") %>%
  set_engine("ranger")

tune_wf <- workflow() %>%
  add_recipe(data_recipe) %>%
  add_model(tune_spec)

set.seed(234)
trees_folds <- vfold_cv(training(data_split), v = 5)

doParallel::registerDoParallel()
set.seed(345)

tune_res <- tune_grid(
  tune_wf,
  resamples = trees_folds,
  grid = 20,
  control = tune::control_grid(verbose = TRUE)
)

tune_res

tune_res %>%
  collect_metrics() %>%
  filter(.metric == "rmse") %>%
  select(mean, min_n, mtry, trees) %>%
  pivot_longer(min_n:trees,
               values_to = "value",
               names_to = "parameter") %>%
  ggplot(aes(value, exp(mean), color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "rmse")

rm(tune_res)
```

```{r}
rf_grid <- grid_regular(
  mtry(range = c(0,37)),
  min_n(range = c(0,28)),
  trees(range = c(0,2000)),
  levels = 10
)

set.seed(456)
regular_res <- tune_grid(
  tune_wf,
  resamples = trees_folds,
  grid = rf_grid,
  control = tune::control_grid(verbose = TRUE)
)

regular_res %>%
  collect_metrics() %>% 
  filter(.metric == "rmse") %>%
  select(mean, min_n, mtry, trees) %>%
  pivot_longer(min_n:trees,
               values_to = "value",
               names_to = "parameter") %>%
  ggplot(aes(value, exp(mean), color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "rmse")

best_auc = select_best(regular_res, "rmse")

final_rf <- finalize_model(
  tune_spec,
  best_auc
)

final_rf

```

```{r}
library(tidymodels) 

set.seed(290) 

n.cores = parallel::detectCores()-1

rf_mod <-
  rand_forest(
    mtry = 4, 
    trees = 222, 
    min_n = 27 ) %>% 
  set_mode("regression") %>%
  set_engine("ranger", importance="permutation", num.threads = n.cores, keep.inbag=TRUE)

wf <-
  workflow() %>%
  add_model(rf_mod)

fit_wf <-
  wf %>%
  add_recipe(data_recipe) %>%
  fit(data = train_data) 

```

```{r}

test_results = fit_wf %>%
  predict(test_data) %>%
  bind_cols(test_data) %>%
  metrics(truth = Diffusive_CH4_Flux_Mean, estimate = .pred)%>%
  select(-".estimator") %>%
  mutate(predict = "test") %>%
  pivot_wider(
    names_from = ".metric",
    values_from = ".estimate"
  ) %>%
  rename("Mean Absolute Error" = mae) %>%
  rename("R-squared" = rsq) %>%
  rename("RSME" = rmse)

train_results = fit_wf %>%
  predict(train_data) %>%
  bind_cols(train_data) %>%
  metrics(truth = Diffusive_CH4_Flux_Mean, estimate = .pred)%>%
  select(-".estimator") %>%
  mutate(predict = "train") %>%
  pivot_wider(
    names_from = ".metric",
    values_from = ".estimate"
  )%>%
  rename("Mean Absolute Error" = mae) %>%
  rename("R-squared" = rsq) %>%
  rename("RSME" = rmse)

combined_results <- bind_rows(test_results, train_results)

library(gt)

table_1 = combined_results %>%
  gt(rowname_col = "predict") %>%
  fmt_number(columns = c(RSME, `R-squared`, `Mean Absolute Error`), decimals = 2)

gtsave(table_1, "table_1.png")
```

```{r, warning=FALSE}
# Separate the mutate step for better debugging
result_data_train_data <- train_data %>%
  mutate(
    prediction = predict(fit_wf, new_data = train_data),
    residual = resid(fit_wf, new_data = train_data)
  ) %>%
  mutate(
    residual = resid(fit_wf, new_data = train_data)
  )

# Extract numeric values from the prediction column
result_data_train_data$prediction <- as.numeric(result_data_train_data$prediction$.pred)

# Separate the mutate step for better debugging
result_data_test_data <- test_data %>%
  mutate(
    prediction = predict(fit_wf, new_data = test_data),
    residual = resid(fit_wf, new_data = test_data)
  ) %>%
  mutate(
    residual = resid(fit_wf, new_data = test_data)
  )

# Extract numeric values from the prediction column
result_data_test_data$prediction <- as.numeric(result_data_test_data$prediction$.pred)


# Combine train and test data
result_data_combined <- bind_rows(
  mutate(result_data_train_data, dataset = "Train"),
  mutate(result_data_test_data, dataset = "Test")
)

jpeg(filename =  "Graph 1.jpg", units = "cm", res = 1000, height = 20, width = 20)

# Plot combined data
result_data_combined %>%
  ggplot(aes(x = Diffusive_CH4_Flux_Mean, y = prediction)) + 
  geom_point(alpha = 0.6, color = 'black') + 
  scale_x_log10(labels = scales::number) +
  scale_y_log10(labels = scales::number) +
  geom_smooth(color = '#0077bb', se = FALSE, method = 'loess', formula = 'y ~ x', linewidth = 1) + 
  geom_abline(color = '#cc3311', slope = 1, intercept = 0, linewidth = 1, linetype = 2) + 
  theme_grey(base_size = 11)+
  theme(text = element_text(family = "Arial")) +
  labs(x = expression(bold(Methane ~ observations) ~ log(mmol ~ m^-2 ~ d^-1)),
       y = expression(bold(Methane ~ predictions) ~ log(mmol ~ m^-2 ~ d^-1))) +
  facet_wrap(~ dataset,
             strip.position = "top",
             scales = "fixed")
dev.off()

```

```{r}
library(DALEXtra)
library(vip)
library(DALEX)  
library(ggplot2)
library(vip)

labeller_vars <- read.csv("/Users/robyn/OneDrive/Documents/1 - Uni of Bristol/Y3 Geography/Dissertation/Data/RiverMethaneFlux-master/RiverMethaneFlux-master/data/processed/variables_names.csv") 

# Set seed for reproducibility
set.seed(222)

rf_explain <- 
  explain_tidymodels(
    model = fit_wf,
    data = data_for_model %>% select(-Diffusive_CH4_Flux_Mean), 
    y = data_for_model$Diffusive_CH4_Flux_Mean,
    label = "ranger"
  )

rf_mod_perf <-  model_performance(rf_explain)

rf_var_imp <- 
  model_parts(
    rf_explain,
    loss_function = loss_root_mean_square
    )

# plot(rf_var_imp, show_boxplots = FALSE, max_vars = 20, title = NULL, subtitle = NULL, bar_width = 10)

loss_df = tibble(variable = rf_var_imp$variable, loss = rf_var_imp$dropout_loss) 
loss_20 <- loss_df |> group_by(variable) |> summarize(loss = mean(loss)) |> arrange(desc(loss)) |> filter(variable != "_baseline_") |> slice(1:20)

library(stringr)

vi_label <- loss_20 %>% filter(variable %in% loss_20$variable) %>% left_join(labeller_vars, by = c("variable" = "var")) %>%  as_tibble() 
# %>% mutate(type = str_replace(type, "Biogeochemical", "Biological")) 

vi_label_mean <- vi_label %>% group_by(variable) %>% summarise(loss=mean(sqrt(loss)), label=first(label), type=first(type)) %>% ungroup() %>% arrange(desc(loss)) %>% mutate(label= factor(label, unique(label)))

library(ggpubr)
library(khroma)

jpeg(filename =  "Graph 2.jpg", units = "cm", res = 1000, height = 20, width = 20)

#plot for the variable imporantance
vi_plot <- ggplot(vi_label, aes(x = loss, y = reorder(label, sqrt(loss), FUN = mean)))+
  stat_summary( aes(fill = type), color= NA, geom="bar", fun="mean", alpha=.8)+
  scale_fill_manual(values=c("#229933", "#4477AA","#EE6677", "#CCBB44", "#BBBBBB", "#66CCEE"), name="Category")+
  theme_classic(base_size = 11)+
  labs(y="", fill="Category",x = "Loss (RSME)")+
  theme(legend.position = c(.78,.15), 
        text = element_text(family = "sans"),
        axis.title.x = element_text(size= 12, ),
        axis.text = element_text(size= 12),
        axis.text.y =ggtext::element_markdown( size= 12), 
        legend.title =element_text(face="bold", size = 12),
        legend.text = element_text(size= 12))+
  coord_cartesian(xlim = c(26, 27.5))

vi_plot

dev.off()
```

```{r}
set.seed(123)
# RF partial dependence plots
rf_pdp <- model_profile(rf_explain, N = NULL)

library(forcats)

labeller_vars2 <- read.csv("/Users/robyn/OneDrive/Documents/1 - Uni of Bristol/Y3 Geography/Dissertation/Data/Modelling/Variable_name2.csv") 

loss_df = tibble(variable = rf_var_imp$variable, loss = rf_var_imp$dropout_loss) 
loss_20 <- loss_df |> group_by(variable) |> summarize(loss = mean(loss)) |> arrange(desc(loss)) |> filter(variable != "_baseline_") |> slice(1:20)

vi_label2 <- loss_20 %>% filter(variable %in% loss_20$variable) %>% left_join(labeller_vars2, by = c("variable" = "var")) %>%  as_tibble() 
# %>% mutate(type = str_replace(type, "Biogeochemical", "Biological")) 

vi_label_mean2 <- vi_label2 %>% group_by(variable) %>% summarise(loss=mean(sqrt(loss)), label=first(label), type=first(type)) %>% ungroup() %>% arrange(desc(loss)) %>% mutate(label= factor(label, unique(label)))

pdp_data_plot_free <- rf_pdp$agr_profiles %>% 
  dplyr::select(variable = `_vname_`, x = `_x_`, y_hat = `_yhat_`) %>%
  right_join(vi_label_mean2, by=c("variable")) %>%
  mutate(label=fct_relevel(label, levels(vi_label_mean2$label))) %>% 
#  group_by(variable) %>% 
#  mutate(x = (x - min(x, na.rm = T))/(max(x, na.rm = T) - min(x, na.rm = T))
#  ) %>% 
  drop_na(y_hat)

# Subset for the variables in pdp_data_plot1
pdp_data_plot1 <- subset(pdp_data_plot_free, variable %in% c("sresp_month", "wetland", "hfi", "runoff_month", "aridity_yr", "prec_yr", "NPP_yr", "urban", "temp_yr", "nee_yr"))

# Subset for the variables in pdp_data_plot2
pdp_data_plot2 <- subset(pdp_data_plot_free, variable %in% c("pyearRS", "peatland_cover", "q_month", "nh4_input", "gw_month", "no3_input", "N_groundwater_agri", "precip_month", "P_point", "trees"))

ggtext::element_markdown()

library(stringr)

pdp_plot1 <- pdp_data_plot1 %>% 
  ggplot(aes(x, y_hat, color=type))+
  geom_line(linewidth=1)+
  scale_color_manual(values=c("#229933", "#4477AA","#EE6677", "#BBBBBB", "#66CCEE", "#CCBB44"), name="Category")+
  facet_wrap(~label, scales = "free", ncol = 2, strip.position = "top") +
  theme_grey()+
  labs(x="Variable change", y= expression(bold(Methane) ~ (mmol ~ m^-2 ~ d^-1))) +
  theme(legend.position = "none", strip.text = ggtext::element_markdown())

pdp_plot2 <- pdp_data_plot2 %>% 
  ggplot(aes(x, y_hat, color=type))+
  geom_line(linewidth=1)+
  scale_color_manual(values=c("#229933", "#4477AA","#CCBB44", "#BBBBBB", "#66CCEE", "#EE6677"), name="Category")+
  facet_wrap(~label, scales = "free", ncol = 2, strip.position = "top") +
  theme_grey()+
  labs(x="Variable change", y= expression(bold(Methane) ~ (mmol ~ m^-2 ~ d^-1))) +
  theme(legend.position = "none", strip.text = ggtext::element_markdown())

free1 = pdp_data_plot_free %>% 
  ggplot(aes(x, y_hat, color=type))+
  geom_line(linewidth=1)+
  scale_color_manual(values=c("#229933", "#4477AA","#CCBB44", "#BBBBBB", "#66CCEE", "#EE6677"), name="Category")+
  facet_wrap(~label, scales = "free", ncol = 2, strip.position = "top") +
  theme_grey()+
  labs(x="scaled variable change", y="methane") +
  theme(legend.position = "none", strip.text = ggtext::element_markdown())

library(patchwork)
library(purrr)

jpeg(filename =  "Graph 3.jpg", units = "cm", res = 1000, height = 20, width = 20)

pdp_plot1

dev.off()

jpeg(filename =  "Graph 4.jpg", units = "cm", res = 1000, height = 20, width = 20)

pdp_plot2

dev.off()

jpeg(filename =  "Graph 4b.jpg", units = "cm", res = 1000, height = 20, width = 20)

free1

dev.off()
```


# test on lm_data

```{r}
lm_data_joined <- flux_df %>% 
  left_join(lm_data, by = c("Site_ID", "Date_end"))

lm_data_joined$Diffusive_CH4_Flux_Mean_nmol_m_s = lm_data_joined$Diffusive_CH4_Flux_Mean

final_lm = lm_data_joined[c(137, 138,  142, 143, 144, 145, 146, 147, 148, 149, 150, 151, 152, 153, 154, 155, 156, 157, 158, 159, 160, 161, 162, 163, 164, 165, 166, 167, 168, 169, 170, 171, 172, 173, 174, 175, 176, 177, 178, 179, 180, 181, 182, 183, 184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205, 206, 207, 208, 209, 210, 211, 212, 213, 214, 215, 216, 217, 218, 219, 220, 221, 222, 223, 224, 225, 226, 227, 228, 229, 230, 231, 232, 233)]

final_lm$Diffusive_CH4_Flux_Mean = final_lm$Diffusive_CH4_Flux_Mean_nmol_m_s * 86400 * 1/1000000

final_lm_2= final_lm[-c(94)]
final_lm_3 <- final_lm_2[-52, ]
final_lm_4 <- final_lm_3[-44, ]
final_lm_5 <- final_lm_4[-58, ]

data_for_model_lm <-  final_lm %>%
  select(-all_of(variables_to_remove)) %>% 
  drop_na()

```


```{r}
lm_results = fit_wf %>%
  predict(data_for_model_lm) %>%
  bind_cols(data_for_model_lm) %>%
  metrics(truth = Diffusive_CH4_Flux_Mean, estimate = .pred)%>%
  select(-".estimator") %>%
  mutate(predict = "Liverpool & Manchester") %>%
  pivot_wider(
    names_from = ".metric",
    values_from = ".estimate"
  ) %>%
  rename("Mean Absolute Error" = mae) %>%
  rename("R-squared" = rsq) %>%
  rename("RSME" = rmse)
  
test_results = fit_wf %>%
  predict(test_data) %>%
  bind_cols(test_data) %>%
  metrics(truth = Diffusive_CH4_Flux_Mean, estimate = .pred)%>%
  select(-".estimator") %>%
  mutate(predict = "Test") %>%
  pivot_wider(
    names_from = ".metric",
    values_from = ".estimate"
  ) %>%
  rename("Mean Absolute Error" = mae) %>%
  rename("R-squared" = rsq) %>%
  rename("RSME" = rmse)

train_results = fit_wf %>%
  predict(train_data) %>%
  bind_cols(train_data) %>%
  metrics(truth = Diffusive_CH4_Flux_Mean, estimate = .pred)%>%
  select(-".estimator") %>%
  mutate(predict = "Train") %>%
  pivot_wider(
    names_from = ".metric",
    values_from = ".estimate"
  )%>%
  rename("Mean Absolute Error" = mae) %>%
  rename("R-squared" = rsq) %>%
  rename("RSME" = rmse)

combined_results2 <- bind_rows(train_results, test_results, lm_results)

library(gt)

table_1a = combined_results2 %>%
  gt(rowname_col = "predict") %>%
  fmt_number(columns = c(RSME, `R-squared`, `Mean Absolute Error`), decimals = 2)

gtsave(table_1a, "table_1a.png")
```

```{r}
# Separate the mutate step for better debugging
result_data <- data_for_model_lm %>%
  mutate(prediction = predict(fit_wf, new_data = data_for_model_lm),
         residual = resid(fit_wf, new_data = data_for_model_lm))

result_data$residual = resid(fit_wf, new_data = data_for_model_lm)

# Extract numeric values from the prediction column
result_data$prediction <- as.numeric(result_data$prediction$.pred)

jpeg(filename =  "Graph 5.jpg", units = "cm", res = 1000, height = 20, width = 20)

# Visualize the data
result_data %>%
  ggplot(aes(x = Diffusive_CH4_Flux_Mean, y = prediction)) + 
  geom_point(alpha = 0.6, color = 'black') + 
#  scale_x_log10(labels = scales::number) +
  scale_y_log10(labels = scales::number) +
  geom_smooth(color = '#0077bb', se = FALSE, method = 'loess', formula = 'y ~ x', linewidth = 1) + 
  geom_abline(color = '#cc3311', slope = 1, intercept = 0, linewidth = 1, linetype = 2) + 
  theme_grey(base_size = 11)+
  theme(text = element_text(family = "sans")) +
  labs(x = expression(bold(Methane ~ observations) ~ (mmol ~ m^-2 ~ d^-1)),
       y = expression(bold(Methane ~ predictions) ~ (mmol ~ m^-2 ~ d^-1)))

dev.off()
```

```{r}
data_for_model |> pull(Diffusive_CH4_Flux_Mean) |> mean()
data_for_model_lm |> pull(Diffusive_CH4_Flux_Mean) |> mean()

data_for_model['centered_ch4'] <- data_for_model$Diffusive_CH4_Flux_Mean - mean(data_for_model$Diffusive_CH4_Flux_Mean)
data_for_model_lm['centered_ch4'] <- data_for_model_lm$Diffusive_CH4_Flux_Mean - mean(data_for_model_lm$Diffusive_CH4_Flux_Mean)

data_for_model |> pull(centered_ch4) |> mean()
data_for_model_lm |> pull(centered_ch4) |> mean()

data_for_model |> pull(centered_ch4) |> min()
data_for_model |> pull(centered_ch4) |> max()

data_for_model_lm |> pull(centered_ch4) |> min()
data_for_model_lm |> pull(centered_ch4) |> max()

data_for_model_cen = data_for_model[-c(1)]
data_for_model_lm_cen = data_for_model_lm[-c(63)]

```

```{r}
set.seed(280)

data_split_cen <- initial_split(data_for_model_cen, prop = 3/4)
train_data_cen <- training(data_split_cen)
test_data_cen  <- testing(data_split_cen)

data_recipe_cen = recipe(centered_ch4 ~., data = train_data_cen)
```

```{r}
set.seed(290) 

n.cores = parallel::detectCores()-1

fit_wf_cen <-
  wf %>%
  add_recipe(data_recipe_cen) %>%
  fit(data = train_data_cen) 

```

```{r}
lm_centred_results = fit_wf_cen %>%
  predict(data_for_model_lm_cen) %>%
  bind_cols(data_for_model_lm_cen) %>%
  metrics(truth = centered_ch4, estimate = .pred)%>%
  select(-".estimator") %>%
  mutate(predict = "Local Urban Methane Flux (centred)") %>%
  pivot_wider(
    names_from = ".metric",
    values_from = ".estimate"
  ) %>%
  rename("Mean Absolute Error" = mae) %>%
  rename("R-squared" = rsq) %>%
  rename("RSME" = rmse)

lm_results = fit_wf %>%
  predict(data_for_model_lm) %>%
  bind_cols(data_for_model_lm) %>%
  metrics(truth = Diffusive_CH4_Flux_Mean, estimate = .pred)%>%
  select(-".estimator") %>%
  mutate(predict = "Local Urban Methane Flux") %>%
  pivot_wider(
    names_from = ".metric",
    values_from = ".estimate"
  ) %>%
  rename("Mean Absolute Error" = mae) %>%
  rename("R-squared" = rsq) %>%
  rename("RSME" = rmse)
  
test_results = fit_wf %>%
  predict(test_data) %>%
  bind_cols(test_data) %>%
  metrics(truth = Diffusive_CH4_Flux_Mean, estimate = .pred)%>%
  select(-".estimator") %>%
  mutate(predict = "Global Urban Methane Flux (testing dataset)") %>%
  pivot_wider(
    names_from = ".metric",
    values_from = ".estimate"
  ) %>%
  rename("Mean Absolute Error" = mae) %>%
  rename("R-squared" = rsq) %>%
  rename("RSME" = rmse)

train_results = fit_wf %>%
  predict(train_data) %>%
  bind_cols(train_data) %>%
  metrics(truth = Diffusive_CH4_Flux_Mean, estimate = .pred)%>%
  select(-".estimator") %>%
  mutate(predict = "Global Urban Methane Flux (training dataset)") %>%
  pivot_wider(
    names_from = ".metric",
    values_from = ".estimate"
  )%>%
  rename("Mean Absolute Error" = mae) %>%
  rename("R-squared" = rsq) %>%
  rename("RSME" = rmse)

combined_results3 <- bind_rows(train_results, test_results, lm_results, lm_centred_results)

library(gt)

table_1b = combined_results3 %>%
  gt(rowname_col = "predict") %>%
  fmt_number(columns = c(RSME, `R-squared`, `Mean Absolute Error`), n_sigfig = 2)


table_1b

gtsave(table_1b, "table_1b.png")
```

```{r}
# Separate the mutate step for better debugging
result_data_cen <- data_for_model_lm_cen %>%
  mutate(prediction = predict(fit_wf_cen, new_data = data_for_model_lm_cen),
         residual = resid(fit_wf_cen, new_data = data_for_model_lm_cen))

result_data_cen$residual = resid(fit_wf_cen, new_data = data_for_model_lm_cen)

# Extract numeric values from the prediction column
result_data_cen$prediction <- as.numeric(result_data_cen$prediction$.pred)

jpeg(filename =  "Graph 6.jpg", units = "cm", res = 1000, height = 20, width = 20)

# Visualize the data
result_data_cen %>%
  ggplot(aes(x = centered_ch4, y = prediction)) + 
  geom_point(alpha = 0.6, color = 'black') + 
  scale_y_log10(labels = scales::number) +
  geom_smooth(color = '#0077bb', se = FALSE, method = 'loess', formula = 'y ~ x', linewidth = 1) + 
#  geom_abline(color = '#cc3311', slope = 1, intercept = 0, linewidth = 1, linetype = 2) + 
  theme_grey(base_size = 11)+
  theme(text = element_text(family = "sans")) +
  labs(x = expression(bold(Methane ~ observations) ~ (mmol ~ m^-2 ~ d^-1)),
       y = expression(bold(Methane ~ predictions) ~ (mmol ~ m^-2 ~ d^-1)))

dev.off()

```